Optimal Symmetric Multimodal Templates and Concatenated Random Forests for Supervised Brain Tumor Segmentation (Simplified) with \textit{ANTsR}

### 2014-04-01 paper.tex

\begin{abstract} 
Manual brain tumor segmentation in MRI is time consuming and
inconsistent both longitudinally and across different raters.
Automated methods have the promise of resolving these issues but are
complicated by the lack of prior knowledge concerning tumor location,
spatial extent, shape, possible displacement of normal tissue, and
intensity signature.  These factors all limit the applicability of
traditional segmentation techniques that have demonstrated good
performance in related tasks such as normal brain tissue segmentation.
The multi-year BRATS challenge seeks to
resolve these issues via high-quality labeled data which may be used
to both guide prior-based segmentation algorithms and also to assess performance in a
blinded, international competition.  In this work, we present the
theoretical and practical background that represents the winning
team's entry into the 2013 BRATs challenge.  This winning approach
combined the features of high-performance methods in both medical
image processing and in machine learning within a new prior-based and automated
tumor segmentation protocol.  This approach employs template-based
deformation models, prior-based segmentation and image geometry
measurements to compute rich, subject-specific multiple modality feature sets which drive a random
forest (RF) prediction model.   The RF model demonstrates excellent
classification performance and also, via variable importance ranking,
identifies key discriminative properties from the rich input feature
set.  Some of the most salient features are those that quantify
asymmetry across different imaging modalities.  We describe how we
employ symmetric multimodal brain templates to derive these
performance critical features and how these features are used within
successive applications of random forest models for increasing levels
of feature refinement, i.e. \textit{concatenated random forests}.  Although the derived 
workflow involves many components, the newly developed \textit{ANTsR} open source software 
package significantly facilitates the proposed methodology.  \textit{ANTsR}
integrates Advanced Normalization Tools (ANTs) with \textit{R}'s
cutting-edge statistical and machine learning techniques and state-of-the-art visualization possibilities.  
The reported algorithmic framework was the top-performing entry in the MICCAI 2013 Multimodal Brain Tumor 
Segmentation challenge and, to our knowledge, is the only entry for both challenge years (2012 and 2013) 
which has been made publicly available.
\end{abstract}

>>>>

\begin{abstract} 
Automated brain tumor segmentation in MRI is complicated by the lack
of prior knowledge concerning tumor location, spatial extent, shape,
possible displacement of normal tissue, and intensity signature.
These factors limit the applicability of techniques with established
performance in related tasks such as normal brain tissue segmentation.
We introduce a template-based framework for computing geometric and
asymmetry-driven feature sets from multiple modality neuroimaging.  We
use these features to drive a supervised whole-brain and tumor segmentation
approach based on random forests.  The asymmetry-related features 
demonstrate excellent discriminative properties within this framework. 
This approach is made possible by symmetric multimodality brain templates which are optimal
in terms of both shape and multivariate intensity.  Further
performance improvements are demonstrated by a successive application
of random forest models for increasing levels of feature refinement to
which we refer as {\it concatenated random forests}.  Although the
derived workflow involves many disparate components, the newly
developed \textit{ANTsR} open source software package significantly facilitates
the proposed methodology.  \textit{ANTsR} is a comprehensive interface between the
popular Advanced Normalization Tools (ANTs) and the \textit{R}
statistical project. \textit{ANTsR} integrates proven image analysis methods
for image registration, segmentation, bias correction, and
template building with \textit{R}'s statistical and machine
learning techniques and state-of-the-art visualization possibilities.  
The reported algorithmic framework was the
top-performing entry in the MICCAI 2013 Multimodal Brain Tumor
Segmentation challenge and, to our knowledge, is the only entry for
both challenge years (2012 \& 2013) which has been made publicly
available.
\end{abstract}
